taxize - taxonomic search and retrieval in R

### 2013-06-10 inst/doc/taxize_withcode_app.tex
fixed figure

\begin{figure}[!ht] 
\centering 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{library}(rgbif)
\hlfunctioncall{library}(ggplot2)

@@

\hlcomment{# Make a map}
p <- \hlfunctioncall{gbifmap}(occurr_list) + \hlfunctioncall{guides}(col = \hlfunctioncall{guide_legend}(title = \hlstring{""}, nrow = 3, 
    byrow = TRUE)) + \hlfunctioncall{theme}(legend.position = \hlstring{"bottom"}, legend.key = \hlfunctioncall{element_blank}()) + 
    \hlfunctioncall{coord_equal}()
p
\end{alltt}
\end{kframe}
\end{knitrout}

\caption{a map...} 
\end{figure}

@@

Suppose we have fuzzy coded trait table with 2 traits with 3 respectively 2 modalities:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
(traits <- \hlfunctioncall{read.table}(header = TRUE, sep = \hlstring{";"}, stringsAsFactors = FALSE, text = \hlstring{"taxon;T1M1;T1M2;T1M3;T2M1;T2M2\textbackslash{}nGammarus sp.;0;0;3;1;3\textbackslash{}nPotamopyrgus antipodarum;1;0;3;1;3\textbackslash{}nCoenagrion sp.;3;0;1;3;1\textbackslash{}nEnallagma cyathigerum;0;3;1;0;3\textbackslash{}nErythromma sp.;0;0;3;3;1\textbackslash{}nBaetis sp.;0;0;0;0;0\textbackslash{}n"}))
\end{alltt}
\begin{verbatim}
                     taxon T1M1 T1M2 T1M3 T2M1 T2M2
1             Gammarus sp.    0    0    3    1    3
2 Potamopyrgus antipodarum    1    0    3    1    3
3           Coenagrion sp.    3    0    1    3    1
4    Enallagma cyathigerum    0    3    1    0    3
5           Erythromma sp.    0    0    3    3    1
6               Baetis sp.    0    0    0    0    0
\end{verbatim}
\end{kframe}
\end{knitrout}

And want to match this to a table with abundances:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
(abundances <- \hlfunctioncall{read.table}(header = TRUE, sep = \hlstring{";"}, stringsAsFactors = FALSE, 
    text = \hlstring{"taxon;abundance;sample\textbackslash{}nGammarus roeseli;5;1\textbackslash{}nGammarus roeseli;6;2\textbackslash{}nGammarus tigrinus;7;1\textbackslash{}nGammarus tigrinus;8;2\textbackslash{}nCoenagrionidae;10;1\textbackslash{}nCoenagrionidae;6;2\textbackslash{}nPotamopyrgus antipodarum;10;1\textbackslash{}nxxxxx;10;2\textbackslash{}n"}))
\end{alltt}
\begin{verbatim}
                     taxon abundance sample
1         Gammarus roeseli         5      1
2         Gammarus roeseli         6      2
3        Gammarus tigrinus         7      1
4        Gammarus tigrinus         8      2
5           Coenagrionidae        10      1
6           Coenagrionidae         6      2
7 Potamopyrgus antipodarum        10      1
8                    xxxxx        10      2
\end{verbatim}
\end{kframe}
\end{knitrout}

First we do some basic data-cleaning and create a looktable, that will link taxa in trait table with taxa in the abundance table.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# first we remove ' sp.' from out trait table:}
traits$taxon_cleaned <- \hlfunctioncall{tolower}(\hlfunctioncall{gsub}(\hlstring{" sp."}, \hlstring{""}, traits$taxon))

@@

The we query the taxonomic hierarchy for both tables, since this is the backbone of this procedure:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{require}(taxize)
traits_classi <- \hlfunctioncall{classification}(\hlfunctioncall{get_uid}(traits$taxon_cleaned))
lookup_classi <- \hlfunctioncall{classification}(\hlfunctioncall{get_uid}(lookup$taxon))
\end{alltt}
\end{kframe}
\end{knitrout}

@@

So our abundance data has two Gammarus species, however trait data is only on Genus level.

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{library}(rgbif)
\hlfunctioncall{library}(ggplot2)

@@

\hlcomment{# Make a map}
p <- \hlfunctioncall{gbifmap}(occurr_list) + \hlfunctioncall{guides}(col = \hlfunctioncall{guide_legend}(title = \hlstring{""}, nrow = 3, 
    byrow = TRUE)) + \hlfunctioncall{theme}(legend.position = \hlstring{"bottom"}, legend.key = \hlfunctioncall{element_blank}()) + 
    \hlfunctioncall{coord_equal}()
p
\end{alltt}
\end{kframe}\begin{figure}[h]

{\centering \includegraphics[width=.9\textwidth]{figure/plot_map}

\caption[A map]{A map\label{fig:mapplot_map}}
\end{figure}

\newpage

@@

Suppose we have fuzzy coded trait table with 2 traits with 3 respectively 2 modalities:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
(traits <- \hlfunctioncall{read.table}(header = TRUE, sep = \hlstring{';'}, stringsAsFactors=FALSE, 
                      text = 'taxon;T1M1;T1M2;T1M3;T2M1;T2M2
Gammarus sp.;0;0;3;1;3
Potamopyrgus antipodarum;1;0;3;1;3
Coenagrion sp.;3;0;1;3;1
Enallagma cyathigerum;0;3;1;0;3
Erythromma sp.;0;0;3;3;1
Baetis sp.;0;0;0;0;0
'))
\end{alltt}
\begin{verbatim}
                     taxon T1M1 T1M2 T1M3 T2M1 T2M2
1             Gammarus sp.    0    0    3    1    3
2 Potamopyrgus antipodarum    1    0    3    1    3
3           Coenagrion sp.    3    0    1    3    1
4    Enallagma cyathigerum    0    3    1    0    3
5           Erythromma sp.    0    0    3    3    1
6               Baetis sp.    0    0    0    0    0
\end{verbatim}
\end{kframe}
\end{knitrout}

And want to match this to a table with abundances:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
(abundances <- \hlfunctioncall{read.table}(header = TRUE, sep = \hlstring{';'}, stringsAsFactors=FALSE, 
                          text = 'taxon;abundance;sample
Gammarus roeseli;5;1
Gammarus roeseli;6;2
Gammarus tigrinus;7;1
Gammarus tigrinus;8;2
Coenagrionidae;10;1
Coenagrionidae;6;2
Potamopyrgus antipodarum;10;1
xxxxx;10;2
'))
\end{alltt}
\begin{verbatim}
                     taxon abundance sample
1         Gammarus roeseli         5      1
2         Gammarus roeseli         6      2
3        Gammarus tigrinus         7      1
4        Gammarus tigrinus         8      2
5           Coenagrionidae        10      1
6           Coenagrionidae         6      2
7 Potamopyrgus antipodarum        10      1
8                    xxxxx        10      2
\end{verbatim}
\end{kframe}
\end{knitrout}

First we do some basic data-cleaning and create a lookup-table, that will link taxa in trait table with the abundance table.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# first we remove ' sp.' from out trait table:}
traits$taxon_cleaned <- \hlfunctioncall{tolower}(\hlfunctioncall{gsub}(\hlstring{" sp."}, \hlstring{""}, traits$taxon))

@@

The we query the taxonomic hierarchy for both tables, this will be the backbone of this procedure:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{require}(taxize)
traits_classi <- \hlfunctioncall{classification}(\hlfunctioncall{get_uid}(traits$taxon_cleaned))
lookup_classi <- \hlfunctioncall{classification}(\hlfunctioncall{get_uid}(lookup$taxon))
\end{alltt}
\end{kframe}
\end{knitrout}

@@

So our abundance data has two Gammarus species, however trait data is only on genus level.
